


```{r}
#load packages
library(librarian)
librarian::shelf(dplyr, ggplot2, lubridate, purrr, sf, dataRetrieval, tigris, tidyverse)


```{r}
# Get all stream gauge stations in California
ca_streams <- read_waterdata_monitoring_location(
  state_name = "California",
  site_type = "Stream"
)

```


```{r}
#create a bounding box to filter data on map

# Build a bounding box for a region of interest, then filter spatially
delta_bbox <- sf::st_bbox(c(xmin = -122.1, ymin = 37.8,
                            xmax = -121.2, ymax = 38.5),
                          crs = sf::st_crs(4326))
delta_sites <- sf::st_filter(ca_streams, sf::st_as_sfc(delta_bbox))

# How many stations?
nrow(delta_sites)

# Preview results (geometry column is retained automatically)
delta_sites |>
  select(monitoring_location_id, monitoring_location_name) |>
  head(10)
```


```{r}
options(tigris_use_cache = TRUE)

# Station of interest (already an sf object from delta_sites)
freeport <- delta_sites |>
  filter(monitoring_location_id == "USGS-11447650")

# Compute map extent: ±0.35° around the station
lon <- as.numeric(sf::st_coordinates(freeport)[1, "X"])
lat <- as.numeric(sf::st_coordinates(freeport)[1, "Y"])
buf <- 0.35
map_xlim <- c(lon - buf, lon + buf)
map_ylim <- c(lat - buf, lat + buf)

# Find counties intersecting the visual extent
vis_rect <- sf::st_sfc(
  sf::st_polygon(list(matrix(c(
    map_xlim[1], map_ylim[1], map_xlim[2], map_ylim[1],
    map_xlim[2], map_ylim[2], map_xlim[1], map_ylim[2],
    map_xlim[1], map_ylim[1]
  ), ncol = 2, byrow = TRUE))),
  crs = sf::st_crs(4326)
)
all_ca <- counties(state = "CA", year = 2022, progress_bar = FALSE)
area_counties <- sf::st_filter(all_ca,
    sf::st_transform(vis_rect, sf::st_crs(all_ca)))

# Water-body polygons for those counties
area_water_sf <- map(area_counties$NAME,
    ~area_water("CA", county = .x, year = 2022)) |>
  bind_rows()

# Extract Freeport station coordinates for the text label (avoids
# geom_sf_text CRS warning)
site_coords <- as.data.frame(sf::st_coordinates(freeport))

ggplot() +
  geom_sf(data = area_counties, fill = "#f5f0e8",
          color = "gray60", linewidth = 0.3) +
  geom_sf(data = area_water_sf, fill = "#a6cee3",
          color = "#6baed6", linewidth = 0.1) +
  geom_sf(data = freeport, color = "#e31a1c",
          size = 4, shape = 21, fill = "#e31a1c") +
  geom_text(data = site_coords,
            aes(x = X, y = Y, label = "Freeport\n(11447650)"),
            nudge_x = 0.05, hjust = 0, size = 2.8) +
  coord_sf(xlim = map_xlim, ylim = map_ylim, expand = FALSE) +
  labs(title = "Sacramento River at Freeport (USGS 11447650)",
       x = NULL,
       y = NULL) +
  theme_light() +
  theme(panel.background = element_rect(fill = "#e8f4f8"))
```


```{r}
# What data does Freeport have?
freeport_avail <- read_waterdata_combined_meta(
  monitoring_location_id = "USGS-11447650"
)

# Which parameters are available, and for how long?
freeport_avail |>
  select(parameter_code, parameter_name, statistic_id, begin, end) |>
  arrange(parameter_code, statistic_id)
```

Option C: Temperature vs. discharge: Retrieve both daily mean temperature (00010) and daily mean suspended sediment concentration (80154) at Freeport. Make a scatter plot. Is there a clear relationship? What might you look at next?


```{r}
freeport_temp_file <- "data/freeport_temp.rds"

if (!file.exists(freeport_temp_file)) {
  message("Fetching daily temperature from USGS...")
  freeport_temp <- read_waterdata_daily(
    monitoring_location_id = "USGS-11447650",  # note "USGS-" prefix
    parameter_code = "00010",  # water temperature
    statistic_id = "00003",  # daily mean
    time = c("2020-01-01", "2023-12-31")
  )
  dir.create("data", showWarnings = FALSE)
  saveRDS(freeport_temp, freeport_temp_file)
} else {
  freeport_temp <- readRDS(freeport_temp_file)
}

glimpse(freeport_temp)

freeport_ss <- read_waterdata_daily(
  monitoring_location_id = "USGS-11447650",
  parameter_code = "80154",  # suspnd sedmnt conc
  statistic_id = "00003",  # daily mean
  time = c("2020-01-01", "2023-12-31")
)
# value column holds the daily mean suspended sediment concentration in mg/L


multi <- read_waterdata_daily(
  monitoring_location_id = "USGS-11447650",
  parameter_code = c("00010", "00060", "80154"),  # temp, discharge, susp sedmnt conc
  statistic_id = "00003",  # daily mean for all three
  time = c("2020-01-01", "2023-12-31")
)

# Plot time series faceted by parameter
multi %>%
  ggplot(aes(x = time, y = value, col = parameter_code)) +
  geom_line() +
  facet_wrap(~parameter_code, scales = "free_y", ncol = 1)




#plot scatter plot
g_temp_sed <- ggplot(multi_2, aes(x = `00010`, y = `80154`)) +
    geom_point()+
    labs(x = "Water Temperature",
    y = "Suspended Sediment Concentration")
g_temp_sed
```


```{r}
#here is an update!
```